Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2018-07-18 ms.tex
Update on Overleaf.

Radiation from massive stars dominates the total feedback energy output of a stellar population \citep[e.g.][]{Lietherer1999,Agertz2013}, surpassing the energy ejection of supernova ($\sim 10^{51}$~erg) by two orders of magnitude. If the radiation couples effectively to the interstellar medium (ISM), it can be a substantial source of additional feedback. %Lower mass stars (size range) primarily drive photoelectric heating (cite) and H$_2$ dissociation through radiation in the far ultraviolet and Lyman-Werner bands, while more massive stars (range) emit significant amounts of ionizing radiation. 
Simulations including stellar radiation feedback followed through radiative transfer or radiation-hydrodynamics schemes have found it to be effective in regulating star formation and driving galactic winds \citep[e.g.][]{Wise2012,Kim2013,Sales2014,Rosdahl2015,Ocvirk2015,Pawlik2015,Peters2017}. This occurs for four reasons: 1) heating gas and preventing the formation of cold, dense star formation regions, 2) the destruction of cold, dense gas around recently formed stars, preventing further star formation, 3) momentum input by direct absorption of UV radiation by gas and (in some cases) dust through re-emission and scattering in the infrared, and 4) lowering the typical ISM densities in which SNe occur and greatly increasing their effectiveness. % the relevance / importance of RP from dust in the IR is beyond the scope of this work...

However, most works that employ stellar radiation feedback to account for these effects do so using various forms of sub-grid, approximate models to avoid the substantial additional cost of full radiative transfer. Many works use a Stromgren approximation whereby the particles / cells within the Stromgren radius a radiating star are heated and kept ionized, with additional approximations made to account for overlapping ionized regions \citep[e.g.][]{HQM2011,Hu2016,Hu2017}. Other works employ some form of energy or momentum injection localized to the region immediately around a star particle \citep[e.g.][(\textit{need more})]{Agertz2013,Roskar2014,Ceverino2014,Forbes2015}. Although some of these approximate methods account for long-range effects of diffuse radiation \citep{HQM2012,Hopkins2018} most cases treat local radiation {\it only}, confined to energy or momentum injection in a limited physical region around a star particle. It is unclear what, if any, effect the long-range ionization and heating processes play in regulating star formation and driving galactic winds. Although the approximate radiation feedback models are capable of driving significant galactic outflows from galaxies, it is unclear if they capture the full effects of stellar ionizing radiation.

@@

\section{Methods and Initial Conditions} \label{sec:methods}
We refer the reader to Paper I for a more detailed description of our methods, briefly summarized here. We use the adaptive mesh refinement hydrodynamics code \textsc{Enzo} \citep{Enzo} to evolve an idealized, isolated low mass dwarf galaxy initialized as a smooth exponential disk set in hydrostatic equillibrium with a static dark matter potential with $M_{\rm gas} = 2 \times 10^6~$~M$_{\odot}$, radial and vertical gas scale heights of 250~pc and 100~pc respectively, $M_{\rm vir} = 2.48 \times 10^9$~M$_{\odot}$, and a maximum physical resolution of 1.8~pc. We include a UV background, tabulated metal line cooling, and a 9 species non-equillibrium chemistry solver using \textsc{Grackle} \citep{Grackle}. Star formation occurs stochastically in cold, dense regions ($n > 200$~cm$^{-3}$, $T < 200$~K) by sampling the \cite{Salpeter1995} IMF from 1~M$_{\odot}$ to 100~M$_{\odot}$ and depositing \textit{individual} star particles over this mass range. For each star, we include feedback from stellar winds, AGB winds, FUV and LW band radiation which drives photoelectric heating and H$_2$ dissociation respectively, HI and HeI ionizing radiation, and core collapse and Type Ia SNe using thermal energy injection. FUV and LW band radiation are both taken to be optically thin, which approximate, local self-shielding alone, but ionizing radiation is followed with radiative transfer (see below).

\textbf{Stellar Ionizing Radiation Models} $-$ We present a comparison of three simulations in this work, which are identical with the exception of their stellar ionizing radiation feedback. In our fiducial simulation, we track the HI and HeI photoionizing radiation using the adaptive ray-tracing radiative transfer method of \citep{MORAY} for all stars above 8$M_{\odot}$ (see Paper I for details on determining the ionizing photon rates for each star). We additionally present a simulation run without any stellar ionizing radiation (``noRT'') and a second simulation to test the relative importance of short-range vs. long-range effects of the radiation where we follow the full ionizing radiation RT, but delete photons once they have traveled more than 20~pc from their source (``shortrad''). This last simulation performs a similar, though significantly more detailed, function as commonly used methods to account for stellar radiation feedback which either deposit thermal energy (citation) or thermal energy and ionization (citations) locally, typically within the Stromgren radius, of each star.  Each simulation is identical up until the formation of the first star particles, at which point they begin to diverge. We note that the stars from the very first star formation event in each run (21 star particles with a total mass of 114 M$_{\odot}$) are the same in each run. Of these, one particle emits ionizing radiation ($M_* > 8 $~M$_{\odot}$).

>>>>

Radiation from massive stars dominates the total feedback energy output of a stellar population \citep[e.g.][]{Lietherer1999,Agertz2013}, surpassing the energy ejection of supernova ($\sim 10^{51}$~erg) by two orders of magnitude. If the radiation couples effectively to the interstellar medium (ISM), it can be a substantial source of additional feedback. %Lower mass stars (size range) primarily drive photoelectric heating (cite) and H$_2$ dissociation through radiation in the far ultraviolet and Lyman-Werner bands, while more massive stars (range) emit significant amounts of ionizing radiation. 
Simulations including stellar radiation feedback followed through radiative transfer or radiation-hydrodynamics schemes have found it to be effective in regulating star formation and driving galactic winds \citep[e.g.][]{WiseAbel2012,Kim2013a,Sales2014,Oshea2015,Rosdahl2015,Ocvirk2015,Pawlik2015,Peters2017}. This occurs for four reasons: 1) heating gas and preventing the formation of cold, dense star formation regions, 2) the destruction of cold, dense gas around recently formed stars, preventing further star formation, 3) momentum input by direct absorption of UV radiation by gas and (in some cases) dust through re-emission and scattering in the infrared, and 4) lowering the typical ISM densities in which SNe occur and greatly increasing their effectiveness. % the relevance / importance of RP from dust in the IR is beyond the scope of this work...

However, most works that employ stellar radiation feedback to account for these effects do so using various forms of sub-grid, approximate models to avoid the substantial additional cost of full radiative transfer. Many works use a Stromgren approximation whereby the particles / cells within the Stromgren radius a radiating star are heated and kept ionized, with additional approximations made to account for overlapping ionized regions \citep[e.g.][]{HQM2011,Hu2016,Hu2017}. Other works employ some form of energy or momentum injection localized to the region immediately around a star particle \citep[e.g.][(\textit{need more})]{Agertz2013,Roskar2014,Ceverino2014,Forbes2016}. Although some of these approximate methods account for long-range effects of diffuse radiation \citep{HQM2012,Hopkins2018} most cases treat local radiation {\it only}, confined to energy or momentum injection in a limited physical region around a star particle. It is unclear what, if any, effect the long-range ionization and heating processes play in regulating star formation and driving galactic winds. Although the approximate radiation feedback models are capable of driving significant galactic outflows from galaxies, it is unclear if they capture the full effects of stellar ionizing radiation.

@@

\section{Methods and Initial Conditions} \label{sec:methods}
We refer the reader to Paper I for a more detailed description of our methods, briefly summarized here. We use the adaptive mesh refinement hydrodynamics code \textsc{Enzo} \citep{Enzo2014} to evolve an idealized, isolated low mass dwarf galaxy initialized as a smooth exponential disk set in hydrostatic equillibrium with a static dark matter potential \citep{Burkert} with $M_{\rm gas} = 2 \times 10^6~$~M$_{\odot}$, radial and vertical gas scale heights of 250~pc and 100~pc respectively, $M_{\rm vir} = 2.48 \times 10^9$~M$_{\odot}$, and a maximum physical resolution of 1.8~pc. We include a UV background, tabulated metal line cooling, and a 9 species non-equillibrium chemistry solver using \textsc{Grackle} \citep{GrackleMethod}. Star formation occurs stochastically in cold, dense regions ($n > 200$~cm$^{-3}$, $T < 200$~K) by sampling the \cite{Salpeter1955} IMF from 1~M$_{\odot}$ to 100~M$_{\odot}$ and depositing \textit{individual} star particles over this mass range. For each star, we include feedback from stellar winds, AGB winds, FUV and LW band radiation which drives photoelectric heating and H$_2$ dissociation respectively, HI and HeI ionizing radiation, and core collapse and Type Ia SNe using thermal energy injection. FUV and LW band radiation are both taken to be optically thin, which approximate, local self-shielding alone, but ionizing radiation is followed with radiative transfer (see below).

%\textbf{Stellar Ionizing Radiation Models} $-$
We present a comparison of three simulations in this work, which are identical with the exception of their stellar ionizing radiation feedback. In our fiducial simulation, we track the HI and HeI photoionizing radiation using the adaptive ray-tracing radiative transfer method of \citep{WiseAbel2011} for all stars above 8$M_{\odot}$ (see Paper I for details on determining the ionizing photon rates for each star). We additionally present a simulation run without any stellar ionizing radiation (``noRT'') and a second simulation to test the relative importance of short-range vs. long-range effects of the radiation where we follow the full ionizing radiation RT, but delete photons once they have traveled more than 20~pc from their source (``shortrad''). This last simulation performs a similar, though significantly more detailed, function as the commonly used methods for approximating the local effects of stellar radiation feedback. Each simulation is identical up until the formation of the first star particles, at which point they begin to diverge. We note that the stars from the very first star formation event in each run (21 star particles with a total mass of 114 M$_{\odot}$) are the same in each run. Of these, one particle emits ionizing radiation ($M_* > 8 $~M$_{\odot}$).
